Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-20 paper/agree.tex
checkpointing; outlined remainder of semantics section
added definition of code contract; other re-orgs and tweakage

This section gives an overview of AGREE's assume-guarantee reasoning
system, abstractly defining the verification conditions that AGREE
checks in order to establish overall system correctness at the model
level.  The discussion then becomes more concrete, showing how the
AGREE contract specification language is used to specify the
cyber-hardened system of \figref{fig:hardened}. It explains the the
top-level specification of the example system, then shifts to the
specifications for the filter and the monitor.  These two
specifications are those that are synthesized to equivalent CakeML by
SPLAT.

@@

\subsection{Verification Conditions}
\input{verification-conditions}

\subsection{System Contract}
\input{sw-contract}

\subsection{Cyber-hardening the System Implementation}
\input{high-assurance-contracts}

>>>>

This section gives an abstract overview of AGREE's compositional
reasoning system, defining the verification conditions that AGREE
checks in order to modularly establish overall system correctness at
the model level.  The discussion then becomes more concrete,
showing---by example---how the AGREE contract language is used to
specify the cyber-hardened system of \figref{fig:hardened}.  After the
system-level contract, discussion shifts to the filter and monitor
component specifications, which are structured to support code
generation.

@@

\subsection{Model of computation}

The AGREE specification language is based on stream concepts, and
operators, from the Lustre language \cite{10.1145/41625.41641}.  Thus
the setting is synchronous dataflow where the inputs, outputs, and
expressions are data streams that comply with the input assumptions.
Contracts are evaluated in dependency order with inputs being
propagated to outputs through all contracts until they stabilize; as
such, the contracts, and thereby the top-level model, must be
acyclic. (An apparent syntactic cycle, where a component is linked
back to itself, may be broken temporally by inserting delay
elements.) \footnote{KLS: some undefined terms here: evaluation of
contracts, stabilization, synchronous step.}

Once the contracts have stabilized, the model takes a synchronous step
to the next input data in the input stream(s).  The semantics do not
model computation or communication delay.\footnote{KLS: how does a
delay element (mentioned above) work then?}  The output of one
contract is seen at the input of any downstream contract in the same
step of the input data stream.
\footnote{KLS: This subsection should give a helpful abstract description of what AG specifications are about, \ie, constraints on IO stream. Needs a rewrite, however.}

\subsection{Verification conditions}
\input{verification-conditions}

\subsection{System contract}
\input{sw-contract}

\subsection{Cyber-hardening the system implementation}
\input{high-assurance-contracts}

### 2022-03-20 paper/example.tex
added definition of code contract; other re-orgs and tweakage

A monitor captures a relation on input data over time and is thus able to reason about temporal properties of that input.
A monitor raises an alert if the specified temporal properties are ever violated.
The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further, that response must come with the request or in the next step after the request.
As with the filter, the system designer provides the policy, and that policy is typically based on existing AGREE specifications in the SW system.

The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened-certificate}.
Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the untrusted AI route planner.
Having passed AGREE verification, the high-assurance components are ready to be synthesized.

The high-assurance components are automatically synthesized by SPLAT from the AGREE specifications to equivalent models in the CakeML language.
The synthesis includes a proof that it captures the meaning of the AGREE specification in CakeML.
In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML code exactly match those defined by the high-assurance component's guarantees. 
From here, CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.

Preserving the input to output relationship of streams between the AGREE contracts and CakeML lifts the AGREE contract verification results to the actual deployed system.
If the contract model verification succeeds, then the meaning of those results hold for the deployed system.
These results, however, are only valid under additional assumptions on the deployed system: contracts for non-synthesized components accurately model their deployed counterparts, an appropriate schedule exists to sequence component execution following the dependent data-flow in the design, and the communication fabric stitching components together works in harmony with the schedule.
Support and automation for these aspects of the design process are discussed in other works.

>>>>

A monitor captures a relation on input data over time and is thus able to reason about temporal properties of that input.
A monitor raises an alert if the specified temporal properties are ever violated.
The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further, that response must come with the request or in the next step after the request.
As with the filter, the system designer provides the policy, and that policy is typically based on existing AADL/AGREE definitions in the SW system.

The AGREE analysis of the cyber-hardened implementation is shown in
\figref{fig:hardened-certificate}. AGREE has generated hihg-level
evidence justifying the claim that the high-assurance components
guarantee the correct behavior of the SW implementation in the
presence of the untrusted AI route planner.  Having passed AGREE
verification, the high-assurance components are ready to be
synthesized.

High-assurance components are automatically synthesized by SPLAT from
the AGREE specifications to equivalent programs in the CakeML
language.  The synthesis toolchain generates a proof that equates the
meaning of the AGREE specification to the behavior of the CakeML
program.  In other words, for any set of input streams that meet the
component's contract assumptions, the output streams produced from the
synthesized CakeML code exactly match those defined by the
high-assurance component's guarantees.  CakeML provides verified
compilation to binaries for several different platforms, meaning that
the resulting binaries exactly preserve the meaning of the original
CakeML code \cite{cakeml}.

Preserving the input to output relationship of streams between the
AGREE contracts and CakeML lifts the AGREE contract verification
results to the actual deployed system.  If the contract model
verification succeeds, then the meaning of those results hold for the
deployed system.  These results, however, are only valid under
additional assumptions on the deployed system:
\begin{itemize}
\item contracts for non-synthesized components accurately model their deployed
counterparts;
\item an appropriate schedule exists to sequence component
  execution following the dependent data-flow in the design; and
\item the communication fabric stitching components together works in harmony
  with the schedule.
\end{itemize}

Support and automation for these aspects of the design process are
discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}

### 2022-03-20 paper/high-assurance-contracts.tex
added definition of code contract; other re-orgs and tweakage

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

As noted previously, the system with the untrusted AI does not pass
AGREE verification.  Transformations in BriefCASE cyber-harden the
system against malicious AI behavior by inserting a high-assurance
filter and monitor between the AI and the WM as shown in
\figref{fig:hardened}.  (In our context \emph{high-assurance} means
that the behaviors of the components are verified to meet their
contracts.)  In the example, the inserted filter protects the WM from
malformed data and the monitor protects the WM from (possibly malicious)
spontaneous or delayed responses.  These high-assurance components are
added, one at a time, by
\begin{itemize}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{itemize}

@@

BriefCASE creates all the needed AADL for the new high-assurance
component and its connections in the system implementation as part of
the transformation.  It also creates a default AGREE specification
that defines everything about the component behavior but the policy
itself, unless that is provided in the dialogue.  The resulting AGREE
specification from the transformation for the filter is shown in
\figref{fig:filter}.  The single guarantee completely defines the
filter output under any input according to the truth value of the
policy.  Such a strong guarantee is needed for synthesis (a weaker
version suffices for AGREE analysis).

The AGREE specification for the monitor is shown in
\figref{fig:monitor}. As can be seen, it is considerably more complex
than the filter specification.  BriefCASE generates everything but the
statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
the filter, the guarantees completely define the output, as required
for synthesis, in terms of the policy and the alert.  The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  Setting \texttt{is\_latched} as true
makes the alert persistent, meaning that once the alert is raised, it
is always raised; otherwise, it is the complement of the policy value
in the current step.

>>>>

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

As previously discussed, transformations carried out in the BriefCASE
interface have created a cyber-hardened system, protecting against
malicious AI behavior, by inserting a high-assurance filter and
monitor between the AI and the WM (see \figref{fig:hardened}). The
filter protects the WM from malformed data and the monitor protects
the WM from (possibly malicious) spontaneous or delayed responses.
These components were added, one at a time, by
\begin{itemize}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{itemize}

@@

BriefCASE creates all the needed AADL for the new high-assurance
component, and its connections in the system implementation, as part
of the transformation.  It also creates a default AGREE specification
that defines everything about the component behavior, leaving only the
component behavior (policy) to be specified.

\subsubsection{Code contracts}

Generally, AGREE specifications \emph{do not} describe the computation
that a component performs. This is entirely by design: AGREE is
intended to reason about component behavior solely at the
specification level. However, the syntax of AGREE specifications
provides enough expressiveness to support the notion of an
AGREE \emph{code contract}: a component contract from which SPLAT can
extract an implementation.

\begin{definition}[Output guarantee]
An \emph{output guarantee} is a stylized guarantee that fully
specifies the data written to an output port. There are three
possibilities according to whether the output port $p$ is
a \konst{data} port, an \konst{event} port, or an \konst{event data}
port:
\[
\begin{array}{ll}
\konst{data}: &  p = \mathit{e} \\
\konst{event}: &  \konst{event} (p) = \mathit{b} \\
\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
\end{array}
\]
\end{definition}

\begin{definition}[Code contract]
A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
if $E \cup G \subset P_c$ and $E$ is a non-empty list of \konst{eq}
statements and $G$ is a set of code guarantees, one for each element
of $O_c$
\end{definition}

The AGREE code contracts for the filter and monitor are shown in
\figref{fig:filter} and \figref{fig:monitor}. In general, the
\konst{eq}-statements of a code contract keep track of
state and provide data values from which to compute values to place on
the outputs.

\subsubsection{Monitor specification}
 For the monitor, BriefCASE generates everything but the
\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
the filter, the guarantees completely define the output, as required
for synthesis, in terms of the policy and the alert.  The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  At configuration
time, \texttt{is\_latched} has been set to \konst{true}, thus making
the alert persistent, meaning that once the alert is raised, it is
always raised; otherwise, it is the complement of the policy value in
the current step.

### 2022-03-20 paper/semantics.tex
added definition of code contract; other re-orgs and tweakage

>>>>

The temporal operators deal with time in more significant ways:
\[
\begin{array}{rcl}
\sem{\konst{pre}(e)}_t & = & \sem{e}_{t-1}, \mathrm{when}\ t > 0 \\
\sem{e_1 \to e_2}_t & = & \itelse{t=0}{\sem{e_1}_0}{\sem{e_2}_t} \\
\sem{\konst{Hist}(b)}_t & = & \forall n \leq t.\; \sem{b}_n
\end{array}
\]

The temporal operators can be used to construct higher-level operators
from PTLTL, such as \konst{Once} and \konst{Since}; they are also used
to manipulate state.

\begin{itemize}
\item \textit{strmStep}.

Given a code contract an implementation will execute the equations in
order, and then compute the values on outputs.

\item \textit{Contract semantics}

\item \textit{How example looks}

\end{itemize}
